---
title: "Figure 5 The Inr strongly contributes to the degree of Pol II pausing"
author: "Wanqing Shao(was@stowers.org) "
output:
  pdf_document:
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancypagestyle{plain}{\pagestyle{fancy}}
---

# Description

To test if initiator (Inr) sequence can contribute to paused Pol II stability, we took a stably paused promoter *dve*, changed its upstream sequence or its upstream sequence together with the Inr to that of a TATA promoter *Act5C*. To further test if Inr can contribute to Pol II pausing independent of TATA box,, we took the stably paused promoter *dve* and replaced its Inr sequence with that of a TATA promoter *Act5C*.

# Enviroment setup

``` {r setup, message=FALSE, error=FALSE}
library(GenomicRanges, warn.conflicts=F)
library(magrittr)
library(Rmisc)

setwd("/data/analysis_code")
options(knitr.figure_dir = 
        "Figure5_Inr_contributes_to_the_degree_of_Pol II_pausing"
        )

source("shared_code/knitr_common.r")
source("shared_code/ggplot_common.r")
source("shared_code/granges_common.r")
source("shared_code/metapeak_common.r")
source("shared_code/sample_common.r")
```

# Analysis

## Pol II signal changes after Inr replacement

```{r inr_replacement, fig.width=7, fig.height=5}
plasmid_annotations <- import("./plasmid_annotation.bed")

get_exo_metapeak <- function(sample, upstream=100, downstream=101, 
                             smooth=NA, endogeneous = F, dps_sample_path=NULL){
	
  gene <- as.character(subset(sample_list, sample_name == sample)$gene)
	chromosome <- as.character(subset(sample_list, sample_name == sample )$chromosome)
	
	if(endogeneous ==F){
	  
	  	sample_path <- load_bigwig(sample)
	  	region <- plasmid_annotations[seqnames(plasmid_annotations) == chromosome & 
	  	                              plasmid_annotations$name == gene] %>%
	  	          resize(., 1, "start")
	    seqlevels(region) <- chromosome
      metapeak <- exo_metapeak(region, sample_path, 
                               upstream=upstream, downstream=downstream, 
                               sample_name=gene, smooth=smooth)
      metapeak$sample <- paste(metapeak$sample_name, metapeak$strand)
      metapeak
      
	}else{
	  
  	  region <- genome_annotations[grep(gene, genome_annotations$name, ignore.case = T)]
  	  seqlevels(region) <- as.character(seqnames(region))
  	  metapeak <- exo_metapeak(region, dps_sample_path, 
  	                           upstream=upstream, downstream=downstream, 
  	                           sample_name=gene, smooth=smooth)
  	  metapeak$sample <- paste(metapeak$sample_name, metapeak$strand)
      metapeak
	 
	}
}


compare_dmso_and_tri <- function(dmso, tri, name,  plotting = T){
  
  plasmid_dmso <- get_exo_metapeak(dmso, upstream=150, downstream = 151)
  plasmid_tri <- get_exo_metapeak(tri, upstream=150, downstream = 151)
  
  plasmid_tri$reads <- plasmid_tri$reads / sum(abs(plasmid_dmso$reads))
  plasmid_dmso$reads <- plasmid_dmso$reads / sum(abs(plasmid_dmso$reads))

  
  plasmid_dmso$sample_name <- paste(name, "Control")
  plasmid_tri$sample_name <- paste(name, "TRI")
  
  if(plotting){
    metapeak <- rbind(plasmid_dmso, plasmid_tri)
    metapeak.p <- subset(metapeak, strand == "+")
    metapeak.n <- subset(metapeak, strand == "-")
  
  plot <- ggplot(metapeak.p, aes(x = tss_distance, y = reads, fill = strand))+ 
          geom_bar(fill="#B23F49", stat="identity") + 
          geom_vline(xintercept =0, linetype = "dotdash")+
          geom_bar(data=metapeak.n, aes(x=tss_distance, y=reads), 
                   fill="#045CA8", stat="identity")+
          xlab("Distance from TSS (bp)")+ ylab("Normalized reads")+
          facet_wrap(facets = "sample_name", ncol =2 )
  print(plot)
  }
  
  dmso_sig <- subset(plasmid_dmso, tss_distance >0 & tss_distance <= 80)$reads %>%
              abs() %>% sum()
  tri_sig <-  subset(plasmid_tri, tss_distance >0 & tss_distance <= 80)$reads%>%
              abs() %>% sum()

  sig_df <- data.frame(condition = c("dmso", "tri"), 
                       paused_polii = c(dmso_sig, tri_sig), 
                       name = name)
  sig_df$paused_pol_norm <- sig_df$paused_polii / sig_df$paused_polii[1]
  sig_df
}


name_list <-c("dve", "act5c_upstream_dve_fusion", "act5c_dve_downstream_fusion", "dve_act5c_inr")

pol_sig_rep1 <- mapply(compare_dmso_and_tri,
       paste0("reporter_dmso_1h_dps_", name_list, "_rpb3_chipnexus_rep1"),
       paste0("reporter_triptolide_1h_dps_", name_list, "_rpb3_chipnexus_rep1"),
       name_list,list(T), SIMPLIFY = F, USE.NAMES =F)  %>% do.call(rbind, .)

pol_sig_rep2 <- mapply(compare_dmso_and_tri,
       paste0("reporter_dmso_1h_dps_", name_list, "_rpb3_chipnexus_rep2"),
       paste0("reporter_triptolide_1h_dps_", name_list, "_rpb3_chipnexus_rep2"),
       name_list, list(F), SIMPLIFY = F, USE.NAMES =F)  %>% do.call(rbind, .)

pol_sig_rep3 <- mapply(compare_dmso_and_tri,
       paste0("reporter_dmso_1h_dps_", name_list[1:3], "_rpb3_chipnexus_rep3"),
       paste0("reporter_triptolide_1h_dps_", name_list[1:3], "_rpb3_chipnexus_rep3"),
       name_list[1:3], list(F), SIMPLIFY = F, USE.NAMES =F)  %>% do.call(rbind, .)

pol_sig_rep4 <- mapply(compare_dmso_and_tri,
       paste0("reporter_dmso_1h_dps_", name_list[1:2], "_rpb3_chipnexus_rep4"),
       paste0("reporter_triptolide_1h_dps_", name_list[1:2], "_rpb3_chipnexus_rep4"),
       name_list[1:2], list(F), SIMPLIFY = F, USE.NAMES =F)  %>% do.call(rbind, .)

process_pol_sig <- function(df, control_n = 1){
  df_sub <- subset(df, condition == "tri")
  df_sub$paused_pol_norm <- df_sub$paused_pol_norm /  df_sub$paused_pol_norm[control_n]
  df_sub
}


sig_df <- lapply(list(pol_sig_rep1, pol_sig_rep2, pol_sig_rep3, pol_sig_rep4),
                process_pol_sig) %>% do.call(rbind, .)

summary_df <- summarySE(sig_df, measurevar="paused_pol_norm", 
                        groupvars=c("name", "condition"))

ggplot(summary_df, aes(x=name, y=paused_pol_norm)) + 
  geom_bar(stat= "identity", position = "dodge", 
          fill =  c("#78AB30", "#3A662F", "#333E2F", "#333E2F")) + 
  geom_errorbar(aes(ymin=paused_pol_norm-se, ymax=paused_pol_norm+se), 
               width=.1, position=position_dodge(.9)) +
  ggtitle("Pol II signal after TRI treatment")+
  ylab("Normalized signal")+
  scale_x_discrete(labels=c("dve", "Act5C-up-dve", 
                           "Act5C-dve-down", "dve-Act5C-Inr")) +
  xlab("")+ 
  geom_hline(yintercept = 1, lty = 4) +
  theme(axis.text.x = element_text(size=14, angle = 60, hjust = 1)) 

```

# Session Info

This analysis was performed with the following R/Bioconductor session:

```{r}
sessionInfo()
```