---
title: "Figure S4 Reporter-ChIP-nexus recapitulates endogenous Pol II pausing"
author: "Wanqing Shao(was@stowers.org) "
output:
  pdf_document:
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancypagestyle{plain}{\pagestyle{fancy}}
---

# Description

To test whether Pol II profile on the plasmid recapitulates the pattern of endogenous promoters, we cloned promoter sequences from *Drosophila pseudoobscura* into our reporter. Pol II profile at *Act5C*, *Pino*, *pepck*, *pxb*, *comm2*, *CG12730*, *pk* and *dve* on the reporter or the endogeneous locus are plotted. Gene-specific 5’ RNA sequencing was used to manually map the transcription start site.

# Enviroment setup

``` {r setup, message=FALSE, error=FALSE}
library(GenomicRanges, warn.conflicts=F)
library(magrittr)

setwd("/data/analysis_code")
options(knitr.figure_dir = 
          "FigureS4_reporter_chipnexus_recapitulates_endogenous_polii_pausing"
        )

source("shared_code/knitr_common.r")
source("shared_code/ggplot_common.r")
source("shared_code/granges_common.r")
source("shared_code/metapeak_common.r")
source("shared_code/sample_common.r")
```

# Analysis

## Pol II profile on the plasmid

```{r plasmid, fig.width = 13}
plasmid_annotations <- import("./plasmid_annotation.bed")
genome_annotations <- import("./dps_genome_annotation.bed")

dps_sample_path <- load_bigwig("genome_dps_dmso_1h_rpb3_chipnexus")

get_exo_metapeak <- function(sample, upstream=100, downstream=101, 
                             smooth=NA, endogeneous = F, dps_sample_path=NULL){
	
  gene <- as.character(subset(sample_list, sample_name == sample)$gene)
	chromosome <- as.character(subset(sample_list, sample_name == sample )$chromosome)
	
	if(endogeneous ==F){
	  
	  	sample_path <- load_bigwig(sample)
	  	region <- plasmid_annotations[seqnames(plasmid_annotations) == chromosome & 
	  	                              plasmid_annotations$name == gene] %>%
	  	          resize(., 1, "start")
	    seqlevels(region) <- chromosome
      metapeak <- exo_metapeak(region, sample_path, 
                               upstream=upstream, downstream=downstream, 
                               sample_name=gene, smooth=smooth)
      metapeak$sample <- paste(metapeak$sample_name, metapeak$strand)
      metapeak
      
	}else{
	  
  	  region <- genome_annotations[grep(gene, genome_annotations$name, ignore.case = T)]
  	  seqlevels(region) <- as.character(seqnames(region))
  	  metapeak <- exo_metapeak(region, dps_sample_path, 
  	                           upstream=upstream, downstream=downstream, 
  	                           sample_name=gene, smooth=smooth)
  	  metapeak$sample <- paste(metapeak$sample_name, metapeak$strand)
      metapeak
	 
	}
}

genome_examples <- c("reporter_dps_act5c_rpb3_chipnexus", 
                      "reporter_dps_pino_rpb3_chipnexus", 
                      "reporter_dps_pepck_rpb3_chipnexus", 
                      "reporter_dps_pxb_rpb3_chipnexus",
                      "reporter_dps_comm2_rpb3_chipnexus",
                      "reporter_dps_cg12730_rpb3_chipnexus", 
                      "reporter_dps_pk_rpb3_chipnexus", 
                      "reporter_dps_dve_rpb3_chipnexus")


genome_metapeaks <- mclapply(genome_examples, function(x){
  metapeak <- get_exo_metapeak(x, upstream=150, downstream = 151,
                               endogeneous = T, dps_sample_path= dps_sample_path)
  metapeak$reads <- metapeak$reads / sum(abs(metapeak$reads))
  metapeak
}, mc.cores =3) %>% do.call(rbind, .)

plot_exo_single_gene <- function(metapeak, name, ncol = 1, scale = "free", tss = 4){
  
  metapeak$sample_name <- factor(metapeak$sample_name, 
                                 levels = unique(metapeak$sample_name))
  
  metapeak.p <- subset(metapeak, strand == "+")
  metapeak.n <- subset(metapeak, strand == "-")

  x <- ggplot(metapeak.p, aes(x=tss_distance, y=reads)) + 
       geom_bar(fill="#B23F49", stat="identity") + 
       geom_bar(data=metapeak.n, aes(x=tss_distance, y=reads),
                fill="#045CA8", stat="identity") + 
       ggtitle(name)+ 
       xlab("Distance from TSS (bp)") + 
       ylab("Normalized reads") + 
       geom_vline(xintercept=0, linetype=tss)+
       facet_wrap(facets = "sample_name", ncol =ncol, scale = scale)
 
    print(x)
}

plot_exo_single_gene(genome_metapeaks, "Endogeneous Pol II profile", 
                     ncol = length(genome_examples), scale = "fixed")
```

## Endogeneous Pol II profile

```{r endogeneous, fig.width = 15}
plasmid_examples <- c(genome_examples, "reporter_scp_rpb3_chipnexus")

plasmid_metapeaks <- mclapply(plasmid_examples, function(x){
  metapeak <- get_exo_metapeak(x, upstream=150, downstream = 151)
  metapeak$reads <- metapeak$reads / sum(abs(metapeak$reads))
  metapeak
}, mc.cores =3) %>% do.call(rbind, .)


plot_exo_single_gene(plasmid_metapeaks, "Pol II profile on the plasmid", 
                     ncol = length(plasmid_examples), scale = "fixed")

```

## Gene-specific 5' RNA sequencing profile

```{r rna_5_sequencing, fig.width=15}
rna_5_samples <- c("reporter_dps_act5c_rna_5_sequencing", 
                   "reporter_dps_pino_rna_5_sequencing",
                   "reporter_dps_pepck_rna_5_sequencing",
                   "reporter_dps_pxb_rna_5_sequencing",
                   "reporter_dps_comm2_rna_5_sequencing",
                   "reporter_dps_cg12730_rna_5_sequencing",
                   "reporter_dps_pk_rna_5_sequencing",
                   "reporter_dps_dve_rna_5_sequencing",
                   "reporter_scp_rna_5_sequencing")

rna_5_metapeaks <- mclapply(rna_5_samples, function(x){
  metapeak <- get_exo_metapeak(x, upstream=150, downstream = 151)
  metapeak$reads <- metapeak$reads / sum(abs(metapeak$reads))
  metapeak
}, mc.cores =3) %>% do.call(rbind, .)

plot_exo_single_gene(rna_5_metapeaks, "Gene-specific 5' RNA sequencin", 
                     ncol = length(plasmid_examples), scale = "fixed", 
                     tss = 0)
```


# Session Info

This analysis was performed with the following R/Bioconductor session:

```{r}
sessionInfo()
```